---
title: "9-25-2018 Class Work"
author: "J Shearston"
date: "September 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```

## Gather

PULSE data import

```{r}
pulse_data = haven::read_sas("./public_pulse_data.sas7bdat") %>%
  janitor::clean_names()
#View(pulse_data)

```

Tidy pulse_data

```{r, warning = FALSE}
pulse_tidy_data = gather(pulse_data, key = visit, value = bdi, bdi_score_bl:bdi_score_12m)
#pulse_tidy_data

```

Illustrate `separate`:

```{r}
pulse_tidy_data %>% 
  separate(visit, into = c("bdi_str", "score_str", "visit"), sep = "_") %>% 
  select(-bdi_str, -score_str) %>% 
  mutate(visit = replace(visit, visit == "bl", "00m"))

```

All together, the data import / cleaning pipeline is:

```{r, warning = FALSE}
pulse_data = haven::read_sas("./public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  gather(key = visit, value = bdi, bdi_score_bl:bdi_score_12m) %>%
  separate(visit, into = c("remove_1", "remove_2", "visit"), sep = "_") %>%
  select(id, visit, everything(), -starts_with("remove")) %>%
  mutate(visit = replace(visit, visit == "bl", "00m"),
         visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)
#pulse_data  
```


## Example: Cleaning the Litters Dataset

Read in the litters dataset

```{r}
read_csv("./FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  count(group) 
```

Use `separate` to tidy up

```{r}
litters_data = 
  read_csv("./FAS_litters.csv", col_types = "ccddiiii") %>% 
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(dose = tolower(dose),
         wt_gain = gd18_weight - gd0_weight) %>%
  arrange(litter_number)
```


## Learning Assessment #1

Selecting variables of interest

```{r}
litters_data %>% 
  select(litter_number, ends_with("weight")) %>%
  gather(key = gd, value = weight, gd0_weight:gd18_weight) %>% 
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))
```


## An example of `Spread`

Creating a "tidy" dataset to illustrate the `spread` command

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)
```

Using the `spread` command

```{r}
spread(analysis_result, key = time, value = mean)
#analysis_result

```


## Binding Rows

Read in datasets

```{r}
fellowship_ring = readxl::read_excel("./LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = readxl::read_excel("./LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = readxl::read_excel("./LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

Bind and tidy

```{r}
lotr_tidy = bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  gather(key = sex, value = words, female:male) %>%
  mutate(race = tolower(race)) %>% 
  select(movie, everything()) 

#lotr_tidy
```


## Joining Datasets

Loading & tidying needed datasets

```{r}
pup_data = read_csv("./FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = read_csv("./FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group))

```

Joining datasets

```{r}
FAS_data = left_join(pup_data, litter_data, by = "litter_number")
#FAS_data

```


## Learning Assessment #2

Importing & tidying class survey data

```{r}
os = read_csv("./surv_os") %>%
  janitor::clean_names()

git_ex = read.csv("./surv_program_git") %>%
  janitor::clean_names()

```

Joining two datasets

```{r}

```

Write a code chunk that imports and cleans both datasets, and then joins them.